Structured Versus Decorated Cospans

### 2021-06-18 cospans_new.tex
Working on the proof of Thm. 2.5 - don't look at it yet.

One can also check that $\Phi$ is an opfibration. Given a morphism $f \maps a \to a'$ in $\A$ and an object $(a \xrightarrow{i}b, x) \in \A/U$ over $a$, we claim that a cocartesian lifting of $f$ is given by 
\[ (f,g,1) \maps (a \xrightarrow{i} b, x) \, \to \, (a' \xrightarrow{i'} b+_a a', F(g)(x)). \]
where $g, i'$ and $b +_a a'$ arise from this pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$b$};
\node (D) at (1.5,-1) {$b+_a a'$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge [dashed] node[right]{$i'$} (D)
(C) edge [dashed] node[below]{$g$} (D);
\end{tikzpicture}
\]
Suppose we are given a morphism 
\[    (f'',g'',h'') \maps (a \xrightarrow{i} b, x) \; \to \; (a'' \xrightarrow{i''} b'', x'')   \]
in $\A/U$ and a morphism $k \maps a \to a''$ in $\A$ such that $\Phi(f'',g'',h'') = k \circ f$.   To show that $(f,g,1)$ is a cocartesian lifting of $f$, we need to show there exists a unique morphism
\[     (\tilde{f}, \tilde{g}, \tilde{h}) \maps (a' \xrightarrow{i'} b +_a a', F(g)(x)) \; \to \; (a'' \xrightarrow{i''} b'', x'') \]
such that $\Phi(\tilde{f}, \tilde{g}, \tilde{h}) = k$ and 
\[       (\tilde{f}, \tilde{g}, \tilde{h}) \circ (f,g,1) = (f'',g'',h'') .\]
To achieve this we choose $\tilde{f} = k$, $\tilde{h} = h''$, and define $\tilde{g}$ using
the universal property of the pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$b$};
\node (D) at (1.5,-1) {$b+_a a'$};
\node (E) at (2.5,-1.5) {$b''$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge node[left]{$i'$} (D)
(C) edge node[above]{$g$} (D)
(C) edge [bend right] node[below]{$g''$} (E)
(B) edge [bend left] node[right]{$i'' k$} (E)
(D) edge [dashed] node[below left]{$\exists! \, \tilde{g} \phantom{a}$} (E);
\end{tikzpicture}
\]
after noting that $g'' i = i' f'' = i'' k f $ by the commuting square condition for the morphisms
in $\A/U$.   One can check that with these choices, $(\tilde{f},\tilde{g}, \tilde{h})$ obeys the desired conditions and is the unique morphism to do so.

@@

\begin{thm}
\label{thm:functoriality}
Given two categories $\A$ and $\A'$ with finite colimits, two lax monoidal pseudofunctors $F \maps \A \to \Cat$ and $F' \maps \A' \to \Cat$, a finite colimit preserving functor $H \maps \A \to \A'$, a lax monoidal pseudofunctor $E \maps \Cat \to \Cat$ and a natural isomorphism $\theta \maps EF \Rightarrow F' H$ as in the following diagram:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\A$};
\node (B) at (1,0) {$\Cat$};
\node (C) at (0,-1) {$\A'$};
\node (D) at (1,-1) {$\Cat$};
\node (E) at (0.5,-0.5) {$\Swarrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
(A) edge node[left]{$H$} (C)
(B) edge node[right]{$E$} (D)
(C) edge node[below]{$F'$} (D);
\end{tikzpicture}
\] 
the triple $(H,E,\theta)$ induces a double functor $\lH \maps F\lCsp \to F'\lCsp$ as defined above.
\end{thm}

\begin{proof}   From the lax monoidal pseudofunctor $F \maps \A \to \Cat$, we obtain a Beck--Chevalley monoidal bifibration $\Phi \maps (\A/U) \to \A$ as in \cref{thm:decorated_cospans_1}.  We do the same for the lax monoidal pseudofunctor $F' \maps \A' \to \Cat$ to obtain a Beck--Chevalley monoidal bifibration $\Phi' \maps (\A'/U') \to \A'$.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(\A/U)$};
\node (B) at (2,0) {$(\A'/U')$};
\node (C) at (0,-1.5) {$\A$};
\node (D) at (2,-1.5) {$\A'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[left]{$\Phi$} (C)
(B) edge node[right]{$\Phi'$} (D)
(C) edge node[above]{$H$} (D);
\end{tikzpicture}
\]
Next we construct a strong monoidal functor $G \maps \A/U \to \A'/U'$  that makes the resulting square commute.  Such a commutating square is a morphism in Shulman's 2-category $\mathscr{MF}^{\textrm{fr}}$ of Beck--Chevalley monoidal bifibrations.  Appying Shulman's 2-functor $\mathbb{F}\mathbf{r} \maps \mathscr{MF}^{\textrm{fr}} \to \mathscr{F}r\mathscr{B}i$ to this morphism, \cite[Theorem 14.11]{Shulman2008}, we obtain the desired double functor
\[ \mathbb{H} \maps  F\lCsp \to F'\lCsp .\]

We described $\A/U$ and its monoidal structure in the proof of \cref{thm:decorated_cospans_1}, 
and of course the case of $\A'/U'$ is analogous.   We define a strong monoidal functor $G \maps \A/U \to \A'/U'$ as follows.  First suppose we are given a object $(a \xrightarrow{i} b, x)$ in $(\A/U)$, where recall $x \in F(x)$.    Then we define
\[  G(a \xrightarrow{i} b, x) = (H(a) \xrightarrow{H(i)} H(b), x') \]
where the object $x' \in F'(H(b))$ is picked out by the composite
\[  1 \xrightarrow{\phi} E(1) \xrightarrow{E(p)} E(F(m)) \xrightarrow{\theta_m} F'(H(m)) \]
so that $U'(\theta_m E(p) \phi)=H(m)$, 
Given a pair of morphisms $(f,\tau)$ that make the following square commute:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$b$};
\node (C) at (0,-1) {$U(x)$};
\node (D) at (1,-1) {$U(y)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[above]{$U(\tau)$} (D);
\end{tikzpicture}
\]
we define $$G(f,\tau)=(H(f),\theta_h E(\tau) \phi).$$
which make the following square in $(\A'/U')$ commute:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$H(a)$};
\node (B) at (3,0) {$H(b)$};
\node (C) at (0,-1) {$U'(\theta_m E(p) \phi)$};
\node (D) at (3,-1) {$U'(\theta_n E(q) \phi))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$H(f)$} (B)
(A) edge node[left]{$$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[above]{$U'(\theta_h E(\tau) \phi)$} (D);
\end{tikzpicture}
\]
where $\tau$ and $\theta_h E(\tau) \phi$ are given respectively by:
\[
\begin{tikzpicture}[scale=1.5]
\node (G) at (-3.5,-0.5) {$1$};
\node (H) at (-2,0) {$F(m)$};
\node (I) at (-2,-1) {$F(n)$};
\node (J) at (-2.5,-0.5) {$\scriptstyle\Downarrow\tau$};
\node (A) at (-0.25,-0.5) {$1$};
\node (L) at (0.5,-0.5) {$E(1)$};
\node (B) at (2,0) {$E(F(m))$};
\node (D) at (2,-1) {$E(F(n))$};
\node (C) at (1.5,-0.5) {$\scriptstyle\Downarrow E(\tau)$};
\node (E) at (4,0) {$F^\prime(H(m))$};
\node (F) at (4,-1) {$F^\prime(H(n))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node [above] {$\phi$} (L)
(G) edge node [above] {$p$} (H)
(G) edge node [below] {$q$} (I)
(H) edge node [right] {$F(h)$} (I)
(B) edge node [above] {$\theta_m$} (E)
(D) edge node [above] {$\theta_n$} (F)
(E) edge node [right] {$F'(H(h))$} (F)
(L) edge node[above]{$E(p)$} (B)
(L) edge node[below]{$E(q)$} (D)
(B) edge node[right]{$E(F(h))$} (D);
\end{tikzpicture}
\]

>>>>

One can also check that $\Phi$ is an opfibration. Given a morphism $f \maps a \to a'$ in $\A$ and an object $(a \xrightarrow{i}b, x) \in \A/U$ over $a$, we claim that a cocartesian lifting of $f$ is given by 
\[ (f,g,1) \maps (a \xrightarrow{i} b, x) \, \to \, (a' \xrightarrow{i'} b+_a a', F(g)(x)). \]
where $g, i'$ and $b +_a a'$ arise from this pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$b$};
\node (D) at (1.5,-1) {$b+_a a'$.};
\node at (1.2,-0.8) {\Large{$\ulcorner$}};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge node[right]{$i'$} (D)
(C) edge node[below]{$g$} (D);
\end{tikzpicture}
\]
Suppose we are given a morphism 
\[    (f'',g'',h'') \maps (a \xrightarrow{i} b, x) \; \to \; (a'' \xrightarrow{i''} b'', x'')   \]
in $\A/U$ and a morphism $k \maps a \to a''$ in $\A$ such that $\Phi(f'',g'',h'') = k \circ f$.   To show that $(f,g,1)$ is a cocartesian lifting of $f$, we need to show there exists a unique morphism
\[     (\tilde{f}, \tilde{g}, \tilde{h}) \maps (a' \xrightarrow{i'} b +_a a', F(g)(x)) \; \to \; (a'' \xrightarrow{i''} b'', x'') \]
such that $\Phi(\tilde{f}, \tilde{g}, \tilde{h}) = k$ and 
\[       (\tilde{f}, \tilde{g}, \tilde{h}) \circ (f,g,1) = (f'',g'',h'') .\]
To achieve this we choose $\tilde{f} = k$, $\tilde{h} = h''$, and define $\tilde{g}$ using
the universal property of the pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$b$};
\node (D) at (1.5,-1) {$b+_a a'$};
\node (E) at (2.5,-1.5) {$b''$};
\node at (1.2,-0.8) {\Large{$\ulcorner$}};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge node[right]{$i'$} (D)
(C) edge node[below]{$g$} (D)
(C) edge [bend right] node[below]{$g''$} (E)
(B) edge [bend left] node[right]{$i'' k$} (E)
(D) edge [dashed] node[below left]{$\exists! \, \tilde{g} \phantom{a}$} (E);
\end{tikzpicture}
\]
after noting that $g'' i = i' f'' = i'' k f $ by the commuting square condition for the morphisms
in $\A/U$.   One can check that with these choices, $(\tilde{f},\tilde{g}, \tilde{h})$ obeys the desired conditions and is the unique morphism to do so.

@@

\begin{thm}
\label{thm:functoriality}
Given two categories $\A$ and $\A'$ with finite colimits, two lax monoidal pseudofunctors $F, F' \maps (\A',+) \to (\Cat,\times)$, a finite colimit preserving functor $H \maps \A \to \A'$, a lax monoidal pseudofunctor $E \maps (\Cat,\times) \to (\Cat,\times)$ and a {\john monoidal???} natural isomorphism $\theta$ as in the following diagram:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\A$};
\node (B) at (1,0) {$\Cat$};
\node (C) at (0,-1) {$\A'$};
\node (D) at (1,-1) {$\Cat$};
\node (E) at (0.5,-0.5) {$\Swarrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
(A) edge node[left]{$H$} (C)
(B) edge node[right]{$E$} (D)
(C) edge node[below]{$F'$} (D);
\end{tikzpicture}
\] 
the triple $(H,E,\theta)$ induces a double functor $\lH \maps F\lCsp \to F'\lCsp$ as defined above.
\end{thm}

\begin{proof}   
From the lax monoidal pseudofunctor $F$ we obtain a Beck--Chevalley monoidal bifibration $\Phi \maps \A/U \to \A$ as in \cref{thm:decorated_cospans_1}.  Similarly the lax monoidal pseudofunctor $F'$ gives a Beck--Chevalley monoidal bifibration $\Phi' \maps \A'/U' \to \A'$.  Next we construct a strong monoidal functor $G \maps \A/U \to \A'/U'$  that makes this square commute:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\A/U$};
\node (B) at (1,0) {$\A'/U'$};
\node (C) at (0,-1) {$\A$};
\node (D) at (1,-1) {$\A'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$G$} (B)
(A) edge node[left]{$\Phi$} (C)
(B) edge node[right]{$\Phi'$} (D)
(C) edge node[below]{$H$} (D);
\end{tikzpicture}
\]
Such a commuting square is a morphism in Shulman's 2-category of Beck--Chevalley monoidal bifibrations.  Appying Shulman's 2-functor $\mathbb{F}\mathbf{r}$ to this morphism \cite[Theorem 14.11]{Shulman2008}, we obtain the desired double functor
\[ \mathbb{H} \maps  F\lCsp \to F'\lCsp .\]

We described $\A/U$ and its monoidal structure in the proof of \cref{thm:decorated_cospans_1}, 
and of course the case of $\A'/U'$ is analogous.   We define the strong monoidal functor $G \maps \A/U \to \A'/U'$ as follows.  First we define $G$ on an object $(a \xrightarrow{i} b, x)$ of $\A/U$.
We can treat $x \in F(b)$ as a morphism $x \maps 1 \to F(b)$.    Then we define
\[  G(a \xrightarrow{i} b, x) := (H(a) \xrightarrow{H(i)} H(b), x') \]
where the object $x' \in F'(H(b))$ is picked out by the composite
\[  1 \xrightarrow{\phi_0} E(1) \xrightarrow{E(x)} E(F(b)) \xrightarrow{\;\theta_b\;} F'(H(b)) .\]
Then we define $G$ on a morphism
\[   (f,g,h) \maps (a \xrightarrow{i} b, x) \; \to \; (a' \xrightarrow{i'} b', x') \]
in $\A/U$.   For this we set
\[   G(f,g,h) = (H(f),H(g),???) \]

where $\tau$ and $\theta_h E(\tau) \phi$ are given respectively by:
\[
\begin{tikzpicture}[scale=1.5]
\node (G) at (-3.5,-0.5) {$1$};
\node (H) at (-2,0) {$F(m)$};
\node (I) at (-2,-1) {$F(n)$};
\node (J) at (-2.5,-0.5) {$\scriptstyle\Downarrow\tau$};
\node (A) at (-0.25,-0.5) {$1$};
\node (L) at (0.5,-0.5) {$E(1)$};
\node (B) at (2,0) {$E(F(m))$};
\node (D) at (2,-1) {$E(F(n))$};
\node (C) at (1.5,-0.5) {$\scriptstyle\Downarrow E(\tau)$};
\node (E) at (4,0) {$F^\prime(H(m))$};
\node (F) at (4,-1) {$F^\prime(H(n))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node [above] {$\phi$} (L)
(G) edge node [above] {$p$} (H)
(G) edge node [below] {$q$} (I)
(H) edge node [right] {$F(h)$} (I)
(B) edge node [above] {$\theta_m$} (E)
(D) edge node [above] {$\theta_n$} (F)
(E) edge node [right] {$F'(H(h))$} (F)
(L) edge node[above]{$E(p)$} (B)
(L) edge node[below]{$E(q)$} (D)
(B) edge node[right]{$E(F(h))$} (D);
\end{tikzpicture}
\]
